Measurement of deeply virtual Compton scattering off Helium-4 with CLAS at Jefferson Lab

### 2021-05-01 eg6-letters/long-paper/prc-dvcshelium.tex
Edits for PRC

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions (FSIs) in the nuclear 
medium and offers a unique opportunity to make a measurement free of them. 
Moreover, with a spin-0 nuclear target, the extraction of the GPD from the DVCS data 
is significantly simplified since a single GPD is involved in the process at leading order. However, the 
measurement of the nuclear DVCS is challenging experimentally and the first attempts by 
the HERMES Collaboration \cite{Airapetian:2009cga} led to controversial conclusions. We present here
in detail the more recent measurements by the CLAS Collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the CLAS nuclear DVCS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

The first measurement of nuclear DVCS was performed by the HERMES 
Collaboration~\cite{Airapetian:2009cga}. This experiment covered an array of 
nuclear targets and looked at the $A$ dependence of the
BSA signal. Their main results, reproduced in Figs. \ref{fig:HERMES1} and \ref{fig:HERMES2}, suffer
from large uncertainties, which makes them consistent with the free proton data and prevents us to
reach strong conclusions about possible nuclear effects. Yet, in the coherent DVCS case a rather
strong effect was expected, leading to a conflict between the HERMES results and theoretical
expectations.

@@

An issue with the HERMES measurement and how it is obtained from data has been raised in
Ref. \cite{Guzey:2003jh} to explain the discrepancy with theoretical expectations.
The main concern is that the DVCS process is not fully detected and the scattered target
is instead reconstructed through a missing mass measurement of the other reaction products. The 
issue with this method is that the detector resolution is not good enough to separate the 
coherent and incoherent channels properly. 
Instead, the results are labeled "coherent enriched" and "incoherent enriched" at low and high 
$-t$, respectively. This label is based on the assumption that the very different behavior of the
cross sections of the two channels in $t$ will lead to a clear differentiation. However, the
results in Fig. \ref{fig:HERMES2} show similar behaviors in both sectors of $t$, which  
challenges this assumption and could explain the tension between theory and experiment.

Altogether, large error bars and the failure to properly separate the coherent and 
incoherent channels have strongly impaired the interpretation of the measurement 
and the conclusions that can be
obtained from it. The CLAS experiment presented here has profited largely from this
result and was designed specifically to solve these two issues of low statistics and exclusivity.

@@

Altogether, CLAS provided a large acceptance for momenta 
starting at 200 MeV. The nuclear DVCS experiment took place from October to December 2009
at an electron beam energy of 6.064 GeV, with the beam intensity varying
between 120 and 150 nA. This beam, on the helium-4 target pressurized between 5 and 6 atm,
corresponds to luminosities in the range of $1$ to $1.2 \times 10^{34}$ cm$^{-2}$s$^{-1}$.
During the experiment, the data acquisition operated at a rate of about 3 kHz with about 70\% live-time
using an inclusive electron trigger.

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV at the beam energy of 6~GeV. Therefore, a specific detector was 
needed to detect them. To design the present setup, inspiration was drawn from the 
BONUS setup that also used a GEM-based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons coming out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
ionization electrons drift toward large radii 
rather than toward the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}. Here a summary of key elements is provided.

>>>>

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions (FSIs) in the nuclear 
medium and offers a unique opportunity to make a measurement free of them. 
Moreover, with a spin-0 nuclear target, the extraction of the GPD from the DVCS data 
is significantly simplified since a single GPD is involved in the process at leading order. However, the 
measurement of the nuclear DVCS is challenging experimentally and the first attempts by 
the HERMES Collaboration \cite{Airapetian:2009cga} to unravel an $A$ dependent nuclear
effect have been unsuccessful. We present here
in detail the more recent measurements by the CLAS Collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the CLAS nuclear DVCS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

The first measurement of nuclear DVCS was performed by the HERMES 
Collaboration~\cite{Airapetian:2009cga}. This experiment covered an array of 
nuclear targets and looked at the $A$ dependence of the
BSA signal. Their main results, reproduced in Figs. \ref{fig:HERMES1} and \ref{fig:HERMES2}, suffer
from large uncertainties, which makes them consistent with the free proton data and prevents us to
reach strong conclusions about possible nuclear effects. Yet, in the coherent DVCS case a rather
strong effect was expected, leading to an apparent conflict between the HERMES results and theoretical
expectations.

@@

A characteristic of the HERMES measurement and how it was obtained from data can however
explain the discrepancy with theoretical expectations~\cite{Guzey:2003jh,Airapetian:2009cga}.
The main point being that the DVCS process is not fully detected and the scattered target
is instead reconstructed through a missing mass measurement of the other reaction products. The 
issue with this method is that the detector resolution is not good enough to separate the 
coherent and incoherent channels properly. 
Thus, the results are labeled "coherent enriched" and "incoherent enriched" at low and high 
$-t$, respectively. This label is based on the assumption that the very different behavior of the
cross sections of the two channels in $t$ will lead to a clear differentiation. However, the
results in Fig. \ref{fig:HERMES2} show similar behaviors in both sectors of $t$, which  
challenges this assumption and could explain the tension between theory and experiment.

Altogether, large error bars and the impossibility to properly separate the coherent and 
incoherent channels have strongly impaired the interpretation of the measurement 
and the conclusions that can be
obtained from it. The CLAS experiment presented here has profited largely from this
result and was designed specifically to solve these two issues of low statistics and exclusivity.

@@

Altogether, CLAS provided a large acceptance for momenta 
starting at 200~MeV. The nuclear DVCS experiment took place from October to December 2009
at an electron beam energy of 6.064 GeV, with the beam intensity varying
between 120 and 150 nA. This beam, on the helium-4 target pressurized between 5 and 6 atm,
corresponds to luminosities in the range of $1$ to $1.2 \times 10^{34}$ cm$^{-2}$s$^{-1}$.
During the experiment, the data acquisition operated at a rate of about 3 kHz with about 70\% live-time
using an inclusive electron trigger.

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200~MeV at the beam energy of 6~GeV. Therefore, a specific detector was 
needed to detect them. To design the present setup, inspiration was drawn from the 
BONUS setup that also used a GEM-based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons coming out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
ionization electrons drift toward large radii 
rather than toward the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}. Here a summary of key elements is provided.
